---
title: "analyze_data"
output: pdf_document
---
Setup and import data. Rename columns cause read.csv renames any numerical columns.
```{r setup, include=FALSE}
library(tidyverse)
library(olsrr)
library(reshape2)
library(broom)
data <- read.csv(file = "../../data/full_data.csv")
data = data %>% 
                  rename(
                    `2007` = X2007,  
                    `2008` = X2008,
                    `2009` = X2009,
                    `2010` = X2010,
                    `2011` = X2011,
                    `2012` = X2012,
                    `2013` = X2013,
                    `2014` = X2014,
                    `2015` = X2015,
                    `2016` = X2016,
                    `2017` = X2017
)
```

Select only columns with vaccination rate data (2007-2017) and use regression to find the rate of change from year to year.
Then plot the rate of change and do a regression on the rate of change against spending per capita, population density, etc. to see if any of these factors has any correlation with rate of change of vaccination rate.
```{r plot1}
data2 <- data %>% select(States, `2007`, `2008`, `2009`, `2010`, `2011`, `2012`, `2013`, `2014`, `2015`, `2016`, `2017`) %>% melt()
data2$variable <- as.numeric(data2$variable)
data2 <- data2 %>% mutate(Year = variable + 2006) %>% select(States, Year, value)

#data3 <- data2 %>% filter(States == "Michigan")
#data2
#ggplot(data=data3, aes(x = Year, y = value)) + geom_point() + geom_smooth(method= "lm", se=F)
#data2

models = data2 %>% group_by(States) %>% do(model = lm(value ~ Year, data = .))
trend <- models %>% tidy(model) %>% select(States, term, estimate, p.value) %>% pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>% rename(Intercept = `estimate_(Intercept)`,
                     Year_coefficient = `estimate_Year`,
                     p.value_Intercept = `p.value_(Intercept)`) %>% arrange(Year_coefficient)
trend

ggplot(data=trend, aes(x=reorder(States, Year_coefficient), y=Year_coefficient)) + geom_point() + theme(axis.text.x = element_text(angle=45, hjust=1)) + xlab("State") + ylab("Change in vaccination rate (%)") + ggtitle("Plot of States' change in vaccination rate")

#ggsave(file="../figures/test.png", width=8, height=4, dpi=500)

all_data <- merge(data, trend, by="States")
coeffModel <- lm(Year_coefficient~spending_per_capita + population_density + Median_income + HighSchool_Plus + Bachelors_Plus + spending_per_pupil + Exemption, data=all_data)
summary(coeffModel)
ols_coll_diag(coeffModel)
```

Regress vaccination rate against many factors. Prints out summary and Tolerance/VIF to verify.
```{r regression against vaccination rate}
model3 <- lm(`2017` ~ spending_per_capita + population_density + Median_income + HighSchool_Plus + Bachelors_Plus + spending_per_pupil + Exemption, data=data)

summary(model3)
ols_coll_diag(model3)


```

Regresses vaccination rate against only spending per capita and % of population with a bachelors' degree.
```{r regression2}
model2 <- lm(`2017` ~ spending_per_capita + Bachelors_plus, data=data)
summary(model2)
ols_coll_diag(model2)
#pairs(data %>% select(spending_per_capita, population_density, Median_income, HighSchool_Plus, Bachelors_Plus, spending_per_pupil))
```